Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-14 nips_2018.tex
editing

Now our goal is to search for programs solving the tasks.  We use the simple search strategy of enumerating programs from
the DSL  in decreasing order of their probability,
and then checking if an enumerated program $p$ assigns positive
probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
into the frontier $\mathcal{F}_x$.

@@

The purpose of the recognition model is to amortize the cost of searching for
programs.  It does this by learning to predict programs which are
probable under $(\mathcal{D},\theta)$ while also assigning high
likelihood for a task according to $\probability[x|p]$.  Concretely,
the recognition model $q$ is a neural network that predicts, for each
task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
a distribution over programs, $\probability[p|\mathcal{D},\theta =
  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
aspect of this framing is that the neural network leverages the
structure of the learned DSL, so it is \emph{not} responsible for
generating programs wholesale.  We share this aspect with
DeepCoder~\cite{balog2016deepcoder} and~\cite{menon2013machine}.

@@

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
In the \system algorithm we infer the DSL from a collection of frontiers.
Intuitively, we want the algorithm to
look at  the frontiers and
generalize beyond them, 
both so the DSL can better express the current solutions,
and  also so that the DSL might expose new abstractions
which will later be used to
discover more programs.

Recall from Sec.~\ref{mathematicalFraming} that we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$.
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{equation}
We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting subexpressions from
programs already in the frontiers.
These subexpressions
are fragments of the original programs, and can introduce new variables (Fig.~\ref{fragmentExample}),
which then become new functions in the DSL.
The idea of storing and reusing
fragments of expressions comes from Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing}.

@@

\section{Symbolic Regression Experiments}\label{regressionSection}
\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
As a simple first example of how our algorithm works, we apply \system
to symbolic regression problems.  For each of these problems, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
200 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  Our recognition model here is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\section{Sequence manipulating programs}\label{sequences}
We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
For both these domains we initially provide the system with a generic
minimal set of programming primitives,
and use a bidirectional GRU~\cite{cho2014learning} for the recognition model.
The initial set of primitives includes routines commonly found in Lisp or Scheme interpreters: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other system successful.
We automatically generated  600 string editing tasks  with 4 input/output examples each (Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
At first, \system cannot find any correct programs for most of the tasks.
It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

>>>>

Now our goal is to search for programs solving the tasks.  We use the simple approach of enumerating programs from
the DSL  in decreasing order of their probability,
and then checking if a program $p$ assigns positive
probability to a task ($\probability[x|p] > 0$); if so, we incorporate $p$ 
into the frontier $\mathcal{F}_x$.

@@

The purpose of the recognition model is to amortize the cost of searching for
programs.  It does this by learning to predict programs which are
probable under $(\mathcal{D},\theta)$ while also assigning high
likelihood to a task according to $\probability[x|p]$.  Concretely,
the recognition model $q$ is a neural network that predicts, for each
task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in
\mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the DSL, this defines
a distribution over programs, $\probability[p|\mathcal{D},\theta =
  q(x)]$.  We abbreviate this distribution as $q(p|x)$.  The crucial
aspect of this framing is that the neural network leverages the
structure of the learned DSL, so it is \emph{not} responsible for
generating programs wholesale.  We share this aspect with
DeepCoder~\cite{balog2016deepcoder} and~\cite{menon2013machine}.

@@

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
%In the \system algorithm we infer the DSL from a collection of frontiers.
Intuitively, we want the algorithm to
look at  the frontiers and
generalize beyond them, 
both so the DSL can better express the current solutions,
and  also so that the DSL might expose new abstractions
which will later be used to
discover more programs.
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{equation}
We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting subexpressions from
programs already in the frontiers.
These subexpressions
are fragments of the original programs, and can introduce new variables (Fig.~\ref{fragmentExample}),
which then become new functions in the DSL.
The idea of storing and reusing
fragments of expressions comes from Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing}.

@@

\section{Symbolic Regression Experiments}\label{regressionSection}
\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
As a simple first example, we apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  The recognition model is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\section{Sequence manipulating programs}\label{sequences}
We apply \system to text editing (Section~\ref{textSection}) and list processing (Section~\ref{listSection}).
For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
and initially provide the system with a generic
set of list processing primitives: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other system successful.
We automatically generated 109 string editing tasks  %% with 4 input/output examples each
(Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
At first, \system cannot find any correct programs for most of the tasks.
It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.
